---
title: "EDA Project"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
### Introduction

This is the 4th project of Udacity's Data Analyst NanoDegree Program. We were given several data sources as options to analyze from. I chose the Arizona's 2016 Presidential Campaign Finance from the Federal Election Commission (FEC) website.

The format of this analysis will be as the following: 

1- I will declare my intentions with a hypothesis (if applicable) 

2- Insert a R snippet/code and run it

3- Declare my findings.

And so on...

First I will begin the analysis by exploring basic statistics about the data set. This will help me see the nature of the data, and whether the data needs cleaning or wrangling. Afterwards, I will explore variable and multivariate relationships, by using the methods I have learned in chapter 4, such as scatter, line, box plots and histograms. This is the basic outline of the analysis, but surely I will find interesting things to talk about along the way. 


### Data Preparation and Munging
```{r echo=FALSE, message=FALSE, warning=FALSE}
# import data and packages
az <- read.csv('P-AZ.csv',sep = ',', row.names= NULL,)

colnames(az) <- c(colnames(az)[-1],"x")
az$x <- NULL

library(ggplot2)
library(plyr)
library(dplyr)
library(gridExtra)
library(memisc)
library(reshape2)
library(gender)
library(stringr)
require(magrittr)
require(zipcode)
require(tmap)
require(glue)
require(githubinstall)


# Install choroplethrZip
#install.packages("devtools")
#library(devtools)
#install_github('arilamstein/choroplethrZip@v1.5.0')

x <- c("ggmap", "rgdal", "rgeos", "maptools", "dplyr", "tidyr", "tmap")
lapply(x, library, character.only = TRUE) # load the required packages



# convert zip code to factor
az$contbr_zip <- factor(az$contbr_zip)
#convert from char to date class
az$proper_date <- as.Date(az$contb_receipt_dt, format = '%d-%B-%y')

```




The structure of the data is as the following The file has 19 variables, and these are the most important ones to for the analysis:

+ Candidate's Name 
+ Contributor's City
+ Contributor's Employer
+ Contribution's Occupation
+ Contribution Amount
+ Contribution's Date


I wish if there was a party and a gender column, I will try to it below. 

```{r }


# Add party col
# Note code template was taken from Udacity Forums

index <- c("Johnson, Gary", "Stein, Jill", "McMullin, Evan")
dindex <- c("Clinton, Hillary Rodham", "Sanders, Bernard", "Lessig, Lawrence",
            "O'Malley, Martin Joseph", "Webb, James Henry Jr.")
rindex <- c('Bush, Jeb', "Carson, Benjamin S."
            , "Christie, Christopher J", "Cruz, Rafael Edward 'Ted'",
            "Fiorina, Carly", "Gilmore, James S III" ,
            "Graham, Lindsey O.", "Huckabee, Mike", 
            "Jindal, Bobby", "Kasich, John R.",
            "Paul, Rand", "Perry, James R. (Rick)",
            "Rubio, Marco", "Trump, Donald J.",
            "Walker, Scott" )
attach(az)
az$party[cand_nm %in% index] <- "independent"
az$party[cand_nm %in% dindex] <- "democrat"
az$party[cand_nm %in% rindex] <- 'republican'
detach(az)

# Convert party to factor
az$party <- factor(az$party)
```

I also would like to add other information such as latitudes and longitudes for map analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}
data(zipcode)
az_loc <- subset(zipcode, state == 'AZ')
az$clean_zip <- substring(az$contbr_zip, 1, 5)
az <- merge(az, az_loc, by.x = 'clean_zip', by.y = 'zip')




```

I would also like to integrate population data by zip-code from the 2010 ZCTA census.



```{r}


# Add Gender col to candidates

# Gender indices
m_index <- c('Bush, Jeb', "Carson, Benjamin S."
            , "Christie, Christopher J", "Cruz, Rafael Edward 'Ted'",
            "Gilmore, James S III" ,
            "Graham, Lindsey O.", "Huckabee, Mike", 
            "Jindal, Bobby", "Kasich, John R.",
            "Paul, Rand", "Perry, James R. (Rick)",
            "Rubio, Marco", "Trump, Donald J.",
            "Walker, Scott", "Sanders, Bernard",
            "Lessig, Lawrence", "O'Malley, Martin Joseph",
            "Webb, James Henry Jr.", "Johnson, Gary",
            "McMullin, Evan"
            )

f_index <- c("Clinton, Hillary Rodham", "Fiorina, Carly", "Stein, Jill" )


head(az)
#simple cand_gender
az$cand_gender <-'a'
attach(az)
az$cand_gender[cand_nm %in% m_index] <- "Male"
az$cand_gender[cand_nm %in% f_index] <- "Female"
detach(az)
# convert cand_gender to factor
az$cand_gender <- factor(az$cand_gender)

az$cand_gender

az$cand_nm %in% m_index

az

```
Now that I added candidates' genders, I'll add the contributors' genders, by using the gender package.
```{r echo=FALSE, message=FALSE, warning=FALSE}
# contributers' genders



# Get first names in a seperate col
az$first_name <- str_split_fixed(az$contbr_nm, ", ", 2)[,2]

# Use gender function
gender_df <- gender(as.character(az$first_name), c(1932, 1998),
countries= "United States")

# Assign gender to contributers in az df
names(gender_df)[1] = "first_name"
names(gender_df)[4] = 'contrib_gender'
gender_df <- unique(gender_df)
az <- merge(az, gender_df[ c("first_name", "contrib_gender")])

# convert contrib_gender to factor
az$contrib_gender <- factor(az$contrib_gender)














```


### Exploratory Data Analysis
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a function that gets  top ten counts
top10 <- function(x){
  y <-table(x)
  y <- sort(y, decreasing = T)
  y <- as.data.frame(y)
  y[1:10,]

  
  
}
  
  
  

# count city occurences in data set
city_tab <- top10(az$contbr_city)


ggplot(city_tab, aes(x, Freq))+
  geom_bar(stat='identity')

# top contribs counts
top_parties <-top10(az$party)

ggplot(top_parties, aes(x, Freq))+
  geom_bar(stat='identity')




city_tab


```

I would like to know how the data is distributed. 


```{r echo=FALSE, message=FALSE, warning=FALSE}

#money disterbution

ggplot(az, aes(contb_receipt_amt))+
  geom_histogram(binwidth = 15)+
  coord_cartesian(xlim=c(0:500))


#money disterbution normalized


ggplot(az, aes(contb_receipt_amt))+
  geom_histogram()+
  scale_x_log10()






ggplot(az, aes(contb_receipt_amt))+
  geom_freqpoly()





```


Date distribution 
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Date disterbution
ggplot(az, aes(proper_date))+
  geom_histogram(binwidth = 20)






```

It appears that we have negative numbers, that goes all to -5400. I believe that it represents refunds, since the most receipt comment is receipt.

I want to find out the amount stats without the refunds.


```{r echo=FALSE, message=FALSE, warning=FALSE}

# most popular bill denomination/   
summary(subset(az$contb_receipt_amt, az$contb_receipt_amt > 0 ))
non_zero_rec <- subset(az$contb_receipt_amt, az$contb_receipt_amt > 0 )
tab <- table(non_zero_rec)
str(tab)


tab[tab>1000]


top_deno<- top10(az$contb_receipt_amt)

ggplot(top_deno, aes(x, Freq))+
  geom_bar(stat='identity')




```

I wonder why the odd numbers such as 19, 27 or even 38.  


Below we will see the number of contributions for each candidate and how they break out: 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Most count of cont.
sort(summary(az$cand_nm), decreasing= T)


# top  cand in terms of money
cand_groups <- group_by(az, cand_nm) 

cand_sum <-summarize(cand_groups, 
                      mean(contb_receipt_amt),
                      n = n())


tot_rec <- aggregate(az$contb_receipt_amt, list(az$cand_nm), sum)


tot_rec <- arrange(tot_rec, desc(x))

tot_rec

# Top in money recieved histogram

ggplot(data= tot_rec,
       aes(reorder(Group.1, -x), x)) +
  geom_bar(stat="identity")
  

# Normalized
ggplot(data= tot_rec,
       aes(reorder(Group.1, -x), x)) +
  geom_bar(stat="identity")+
  scale_y_log10()



```

I would like to see the box-plot of each gender/party contribution.
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(az[!is.na(az$party) & az$party != 'independent',],
       aes(x=contrib_gender, y=contb_receipt_amt))+
  geom_boxplot(varwidth=T)+
  facet_grid(~party)+
  coord_cartesian(ylim=boxplot.stats(az$contb_receipt_amt)$stats[c(1, 5)]
)




```
Republicans contributed more on average, and they had a higher range of contribution amounts. Male republicans contributed slightly more on average than their female counterparts. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
table(az$contrib_gender)
```
I did not expect to find more female contributors than males in this data-set. 

Lets explore if females were more likely to vote for female candidates.
```{r echo=FALSE, message=FALSE, warning=FALSE}
same_sexf<- subset(az, az$contrib_gender == 'female' &
                     az$cand_gender== 'Female')

length(same_sexf$contrib_gender)/
  length(az$contrib_gender[az$contrib_gender == 'female'])


same_sexm<- subset(az, az$contrib_gender == 'male' & az$cand_gender== 'Male')

length(same_sexm$contrib_gender)/
  length(az$contrib_gender[az$contrib_gender == 'male'])


ggplot(az[!is.na(az$contrib_gender) & !is.na(az$cand_gender),] ,
       aes(cand_gender, contrib_gender))+
  geom_bin2d()


```
%54.5 females of this data-set contributed to females, while %65.4 of males contributed to males, which is a negligible preference. 



```{r echo=FALSE, message=FALSE, warning=FALSE}

# Create the mode function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

# get distinct empoyee contrib from each employer

employer<- summary(az$contbr_employer)



                                  
# most genrous occupations and account of total 
                                  

#subset academia employees                               
academia <- subset(az, grepl('UNIVERSITY|COLLEGE', az$contbr_employer,
                             ignore.case = T))


# group by university/college
cand_acad <- group_by(academia, contbr_employer)

# how much is averagly contributed by academia
spend_acad <- summarise(cand_acad,
                        number_of_contributions = n(),
                        avg_spent = mean(contb_receipt_amt),
                        main_cand = getmode(cand_nm),
                        party = getmode(party)
                        
                        )

# sort by num of contribs
attach(spend_acad)
spend_acad <- spend_acad[order(-number_of_contributions),]
detach(spend_acad)
spend_acad

# percentage of demo/repub colleges
percent(spend_acad$party)

# Percentage of candidate preferance
sort(percent(academia$cand_nm)[1:24],decreasing = T)

# 

```
Around %80 of colleges had a democratic preference. The majority of %59 of contributions were for Clinton, Sanders cones in second of %37. Cruz came in third (%1.42) and Trump close fourth (%1.4).


Below I will find the stats of homemakers and retirees
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Homemaker stats
homemaker <- subset(az, grepl('HOMEMAKER', az$contbr_occupation, ignore.case = T ))
summary(homemaker)

# Percentage of homemaker genders
percent(homemaker$contrib_gender)

# Homemaker contribs party leaning
percent(homemaker$party)





# retired findings
retired <- subset(az, grepl('RETIRED', az$contbr_occupation, ignore.case = T))

#percentage of retirees in the data
print(
length(unique(retired$contbr_nm))/
  length(unique(az$contbr_nm)) * 100)



# Retired party leaning percentage
print (summary(retired$party)/ length(retired$party) *100)

# Retrired avg spending
mean(retired$contb_receipt_amt)










```
Homemakers are %96 females, and  %65 of homemakers are democrats.

As we can see above, retirees make up about %35.6 of the data-set. Around %60 of retirees contributed to democrats and around %40 percent to republicans, contributions to independents are negligible. Retirees contributed $81 on average.



I want to know which occupations are most politically active, and how do they lean politically.

```{r echo=FALSE, message=FALSE, warning=FALSE}
jobs <-data.matrix(summary(az$contbr_occupation))

head(jobs,10)
```
The most politically active occupations in the data set are attorneys, physicians then teachers.

Below I would like to know the proportions of party leaning for each job. For example, of all engineers how many percent of them lean republican (number of republican engineers/ total number of engineers).
```{r echo=FALSE, message=FALSE, warning=FALSE}
require(GGally)
require(gmodels)


# Create a DF for jobs and their party count
job_party_table<- as.data.frame.matrix(table(az$contbr_occupation, az$party))

# Create a DF with jobs and thier party percentages
job_party_dist<- prop.table(table(az$contbr_occupation, az$party), 1)

job_party_dist <- as.data.frame.table(job_party_dist)

# Merge the two DFs above
job_party_props<- data.frame(c(job_party_table, job_party_dist))

#Keep these cols
keep <- c('Var1','democrat', 'republican', 'Var2', 'Freq')
job_party_props <- job_party_props[,keep]

#change props to percentages
job_party_props$Freq <- job_party_props$Freq *100
# add a total_job col
job_party_props$total <- job_party_props$democrat + job_party_props$republican

#drop these cols
job_party_props$democrat <- NULL
job_party_props$republican <- NULL

# change Var2 col to party
names(job_party_props)[names(job_party_props) == 'Var2'] <- 'party'


# subset top 10 republican jobs
top10rep <-subset(job_party_props, party == 'republican' & total > 1000)

# create a pie chart
pie(top10rep$Freq, labels = top10rep$Var1)

#create a bar plot
ggplot(top10rep,
       aes(Var1, Freq))+
  geom_bar(stat = "identity")



# subset top 10 democrats jobs
top10demo <-subset(job_party_props, party == 'democrat' & total > 1000
                   & quantile(job_party_props$Freq, c(.660), na.rm = T))

# create a bar plot
ggplot(top10demo,
       aes(Var1, Freq))+
  geom_bar(stat = "identity")


# subset top 10 independents jobs
top10inde <-subset(job_party_props, party == 'independent' & 
                     Freq > quantile(Freq, probs = .62, na.rm = T)
                   & total >100 & Var1 != 'N/A')

# create a bar plot
ggplot(top10inde,
       aes(Var1, Freq))+
  geom_bar(stat = "identity")



```






I would like to see average spending along dates
```{r echo=FALSE, message=FALSE, warning=FALSE}
#group by avg week
require(lubridate)

by_wk_sum <- tapply(az$contb_receipt_amt, week(az$proper_date), sum)

by_wk_avg <- tapply(az$contb_receipt_amt, week(az$proper_date), mean)


plot(by_wk_avg, type = 'l')

plot(by_wk_sum)



# note the weeks are aggregated by all years
az$week <-format(az$proper_date, format = "%W")
az$month <-format(az$proper_date, format = "%m")
az$year <-format(az$proper_date, format = "%y")

by_wk <- az %>% group_by(year = as.numeric(year), week= as.numeric(week), party)
%>% summarise(sum = sum(contb_receipt_amt), 
              avg = mean(contb_receipt_amt),
              n=n())



ggplot(subset(by_wk, avg> 0 & year > 14),
       aes(as.numeric(week), avg, color=party, size=n))+ geom_line()+
  facet_grid(~year)


```

It seems that the avg amount of contributions are huge at the beginning of 2015, but when I added a 4th variable (n = number of contributions) it shows that these were a few outliers, the mass of the contributions came in mid 2016 as it lowered the average but the size (n) was bigger substantially. 




```{r echo=FALSE, message=FALSE, warning=FALSE}
library(choroplethrZip)
library(choroplethr)
require(glue)
data(df_zip_demographics)
data("zip.regions")

zip.regions
az_demographics <- subset(zip.regions, region > 85001 & region < 86556) 

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# create a function that selects zipcode and region for zip_choropleth analysis

clean <- function(x, y){
  region <- x
  value <- y
  new <- data.frame(region, value)
 new <- new[!duplicated(new$region), ]
}




# group by other intersting variables.
zip_avg_sum <- az %>%
  group_by(clean_zip, party) %>%
  summarise(
            average_contrib = mean(contb_receipt_amt),
            sum_of_contribs = sum(contb_receipt_amt))


# Democratic contrib heatmap
demo_zip_avg_sum <- na.omit(zip_avg_sum[zip_avg_sum$party == 'democrat',])

demo_zip_avg_sum <- clean(demo_zip_avg_sum$clean_zip, 
                          demo_zip_avg_sum$sum_of_contribs)

demo_heat <- zip_choropleth(demo_zip_avg_sum,state_zoom = 'arizona',
                            county_zoom = 4013 )+
  ggtitle('Maricopa County Total USD Contribution (by Zipcode)', 
          subtitle = 'Democratic Party Candidates')+
  labs(color= 'USD Amount')

# Republican contrib heatmap 
repub_zip_avg_sum <- na.omit(zip_avg_sum[zip_avg_sum$party == 'republican',])
repub_zip_avg_sum <- clean(repub_zip_avg_sum$clean_zip,
                           repub_zip_avg_sum$sum_of_contribs)
repub_zip_avg_sum
repub_heat <- zip_choropleth(repub_zip_avg_sum,
                             state_zoom = 'arizona', county_zoom = 4013)+
  ggtitle(' ',subtitle =  'Republican Party Candidates') + 
  scale_fill_brewer(palette="Reds")
  
# why I can't change the color to red.

repub_heat

# problem: heatmap scales are not matching, 
grid.arrange(demo_heat, repub_heat)
 # Reasearch heatmap color and scale


  
  


```













```{r echo=FALSE, message=FALSE, warning=FALSE}
# Final plot 3
require(choroplethrMaps)

df_zip_demographics <- subset(df_zip_demographics, region > 85001 & region < 86556) 

ALLdemographics_df<- merge.data.frame(df_zip_demographics, zip.regions,
                                      by.x='region', by.y='region')

ALLdemographics_df



az_county<- merge.data.frame(az,ALLdemographics_df,
                             by.x='clean_zip', by.y = 'region')

az_county



az_county_summ <- az_county %>%
group_by(county.name, county.fips.numeric, party) %>%
  summarise(n= n(),
             average_contrib = mean(contb_receipt_amt),
            sum_of_contribs = sum(contb_receipt_amt))



# Clean  avg_contribution
az_countyP1 <- with(az_county_summ, clean(county.fips.numeric, average_contrib))
az_countyP1 <- na.omit(az_countyP1)

# Plot avg_contribution by county
county_choropleth(az_countyP1, state_zoom = 'arizona')

# group by county & party
az_county_party <- merge(x = az_county, y = az[, c("party","tran_id" )],
                         by = "tran_id", all.x=TRUE)

az_county_party_grouped <- az_county_party %>%
  group_by(county.fips.numeric, party.x) %>%
  summarize(average = mean(contb_receipt_amt))

# Plot demo avg_contribution by county

az_countyP1Demo <- 
  with(az_county_party_grouped[az_county_party_grouped$party.x == 'democrat',],
       clean(county.fips.numeric, average))

P1Demo <-county_choropleth(na.omit(az_countyP1Demo), state_zoom = 'arizona')+
    ggtitle(label = 'Democrate Average Contribution', subtitle = 'By County')







# Plot repub avg_contribution by county

az_countyP1Repub <-
  with(az_county_party_grouped[az_county_party_grouped$party.x == 'republican',],
       clean(county.fips.numeric, average))

P1Repub <- county_choropleth(na.omit(az_countyP1Repub), state_zoom = 'arizona')+
  ggtitle(label = 'Republican Average Contribution', subtitle = 'By County') +
  scale_fill_brewer(palette="Reds")



# Arrange the party plots 
grid.arrange(P1Demo, P1Repub)

```
Note: there is discrepancy in the color scale: 





Brain storming:
What can I do to improve?


What kind of graphs could I add?
-Bar chart of most contributing jobs to the Donald

I am wondering what kind jobs contributed to Donald trump, my intuition says it's mostly blue collar jobs. Let's find out!


```{r echo=FALSE, message=FALSE, warning=FALSE}
#-Bar chart of most contributing jobs to the donald

Trump_jobs<- az[az$cand_nm == "Trump, Donald J.",] %>%
group_by(job=contbr_occupation) %>%
  summarise(avg = mean(contb_receipt_amt),
            n = n())

ggplot(Trump_jobs[Trump_jobs$n >100,], aes(x=job, y=avg))+
geom_bar(stat = 'identity')

#-Bar chart of most contributing jobs to Clinton
Clinton_jobs<- az[az$cand_nm == "Clinton, Hillary Rodham",] %>%
group_by(job=contbr_occupation) %>%
  summarise(avg = mean(contb_receipt_amt),
            n = n())

ggplot(Clinton_jobs[Clinton_jobs$n >390 & Clinton_jobs$avg > 0 ,],
       aes(x=job, y=avg))+
geom_bar(stat = 'identity')

unique(az$cand_nm)

```
My hypothesis is false, most of trumps contributors have white collar jobs, even the higher income types such as engineers, consultants, physicians and CEOs. One weakness of this plot, it does not represent low income contributors.

let me see by number of contributions only if it helps me find out more,
```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(Trump_jobs[Trump_jobs$n >100 & Trump_jobs$n < 2000,], aes(x=job, y=n))+
geom_bar(stat = 'identity')

sum(Trump_jobs$n)/sum(cand_sum$n)



ggplot(Trump_jobs[Trump_jobs$n >20 & Trump_jobs$n < 100,], aes(x=job, y=n))+
geom_bar(stat = 'identity')+
   theme(axis.text.x=element_text(angle=90,hjust=1))





```
The percentage of contributions for trump of the whole data-set is 18%.

By changing some of the subset filters, still the majority were high income occupations, even though we have some blue collar jobs such as truck driver and construction, but they were the minority. My hypothesis is blue-collar workers cannot afford to contribute therefore, they are underrepresented in this data-set.


I want to see if higher income zip-codes had more contributions and I will use a scatter-plot to demonstrate.
```{r echo=FALSE, message=FALSE, warning=FALSE}
# final plots 2
# merge demographic data with original dataset
AZ_ALLdemographics_df <- merge.data.frame(az, ALLdemographics_df,
                                          by.x = 'clean_zip', by.y = 'region')
income_cotrib <- AZ_ALLdemographics_df %>%
  group_by(clean_zip, income= per_capita_income)%>% 
  summarise(n = n(), total= sum(contb_receipt_amt),
            average = mean(contb_receipt_amt))



ggplot(income_cotrib[income_cotrib$n >100,], aes(income, total))+
  geom_point()+
  geom_smooth()+
  xlab('Median Income per Zipcode')+
  ylab('Total Population per Zipcoe')+
  ggtitle('Relationship Between Number of Median Income and Population per Zipcode')



ggplot(income_cotrib[income_cotrib$n >100,], aes(income, average))+
  geom_point()+
  geom_smooth()+
  xlab('Median Income per Zipcode')+
  ylab('Average Contibution per Zipcode')+
  ggtitle('Relationship Between Average USD Amount of Contributions and Median Income',
          subtitle = 'Per Zipcode')




```
There is only a strong relationship When I subsetted the data to 100 contributions at least per zip-code. Doing otherwise will skew the data and the relationship will not be apparent. 





This is obvious but still I would like to see the relationship between number of contributions and population of zip-code.
```{r echo=FALSE, message=FALSE, warning=FALSE}
#final plot 1

AZ_ALLdemographics_df

pop_contrib <- AZ_ALLdemographics_df %>%
  group_by(clean_zip, total_population)%>%
  summarize(n=n())

ggplot(pop_contrib, aes(n,total_population))+
  geom_point()+
  geom_smooth()+
  xlim(0,1000)+
  xlab('Number of Contributions')+
  ylab('Total Population')+
  ggtitle('Relationship Between Number of Contributions and Population of per Zipcode')





```

There is a strong correlation at first, but then as population increases the relationship weakens.



### Final Plots:
```{r echo=FALSE, message=FALSE, warning=FALSE}
finalp1<- ggplot(subset(na.omit(by_wk), avg> 0 & year > 14),
       aes(as.numeric(week), avg, color=party, size=n))+ geom_line()+
  facet_grid(~year)+
  ggtitle('Average USD Contributed Along 2015/16 Weeks by Party')+
  xlab('Week #')+
  ylab('USD Contributed (Average)')+
  labs(color= 'Party', size = 'Number of Contributions')
  
finalp1


finalp2 <-  ggplot(pop_contrib, aes(n,total_population))+
  geom_point()+
  geom_smooth()+
  xlim(0,1000)+
  xlab('Number of Contributions')+
  ylab('Total Population')+
  ggtitle('Relationship Between Number of Contributions and Population of per Zipcode')

finalp2

finalp3 <- ggplot(income_cotrib[income_cotrib$n >100,], aes(income, total))+
  geom_point()+
  geom_smooth()+
  xlab('Median Income per Zipcode')+
  ylab('Total Population per Zipcoe')+
  ggtitle('Relationship Between Number of Median Income and Population per Zipcode')

finalp3

finalp4 <- ggplot(income_cotrib[income_cotrib$n >100,], aes(income, average))+
  geom_point()+
  geom_smooth()+
  xlab('Median Income per Zipcode')+
  ylab('Average Contibution per Zipcode')+
  ggtitle('Relationship Between Average USD Amount of Contributions and Median Income',
          subtitle = 'Per Zipcode')

finalp4


finalp5 <-  county_choropleth(az_countyP1, state_zoom = 'arizona')



finalp5 

finalp6 <-  grid.arrange(P1Demo, P1Repub)

finalp6



finalp7 <- grid.arrange(demo_heat, repub_heat)







```



### Reflection:
overall this project was a good challenge and learning experience. At first it was easy and enjoyable exploring the data, as I went deeper into the analysis it became harder to come up with relationships and conclusions about the data. I wanted my analysis to have a central  theme/thesis, the fact of not drawing a certain conclusion made me feel frustrated. 

I was impressed with the versatility of R, and its packages, I felt like it was more intuitive than python, maybe because I have a background with Alteryx. Although, R felt like it had less support on stackoverflow than python, but there's support nonetheless, which aided me significantly throughout the project. I also used Datacamp for filling in the knowledge gaps and reinforcing the concepts learned in the Udacity curriculum. I have not utilized Udacity's live help as much as the other projects, because I did not face problems with programming itself, rather than loss of ideas and direction of my analysis. 

In terms of visualizations, R is fantastic for data exploration, although it is lacking the ability to export high resolution plots. I feel that Tableau is more suitable for findings/conclusive plots. 





