---
title: "EDA Project"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
### Introduction

This is the 4th project of Udacity's Data Analyst NanoDegree Program. We were given several data sources as options to analyze from. I chose the Arizona's 2016 Presidential Campaign Finance from the Federal Election Commission (FEC) website.

The format of this analysis will be as the following: 

1- I will declare my intentions with a hypothesis (if applicable) 

2- Insert a R snippet/code and run it

3- Declare my findings.

And so on...

First I will begin the analysis by exploring basic statistics about the data set. This will help me see the nature of the data, and whether the data needs cleaning or wrangling. Afterwards, I will explore variable and multivariate relationships, by using the methods I have learned in chapter 4, such as scatter, line, box plots and histograms. This is the basic outline of the analysis, but surely I will find interesting things to talk about along the way. 


### Data Preparation and Munging
```{r echo=FALSE, message=FALSE, warning=FALSE}
# import data and packages
az <- read.csv('P-AZ.csv',sep = ',', row.names= NULL,)

colnames(az) <- c(colnames(az)[-1],"x")
az$x <- NULL

library(ggplot2)
library(plyr)
library(dplyr)
library(gridExtra)
library(memisc)
library(reshape2)
library(gender)
library(stringr)
require(magrittr)
require(zipcode)
require(tmap)
require(glue)
require(githubinstall)


# Install choroplethrZip
#install.packages("devtools")
#library(devtools)
#install_github('arilamstein/choroplethrZip@v1.5.0')

x <- c("ggmap", "rgdal", "rgeos", "maptools", "dplyr", "tidyr", "tmap")
lapply(x, library, character.only = TRUE) # load the required packages



# convert zip code to factor
az$contbr_zip <- factor(az$contbr_zip)
#convert from char to date class
az$proper_date <- as.Date(az$contb_receipt_dt, format = '%d-%B-%y')

```




The structure of the data is as the following The file has 19 variables, and these are the most important ones to for the analysis:

+ Candidate's Name 
+ Contributor's City
+ Contributor's Employer
+ Contribution's Occupation
+ Contribution Amount
+ Contribution's Date


I wish if there was a party and a gender column, I will try to it below. 

```{r echo=FALSE, message=FALSE, warning=FALSE}


# Add party col
# Note code template was taken from Udacity Forums

index <- c("Johnson, Gary", "Stein, Jill", "McMullin, Evan")
dindex <- c("Clinton, Hillary Rodham", "Sanders, Bernard", "Lessig, Lawrence",
            "O'Malley, Martin Joseph", "Webb, James Henry Jr.")
rindex <- c('Bush, Jeb', "Carson, Benjamin S."
            , "Christie, Christopher J", "Cruz, Rafael Edward 'Ted'",
            "Fiorina, Carly", "Gilmore, James S III" ,
            "Graham, Lindsey O.", "Huckabee, Mike", 
            "Jindal, Bobby", "Kasich, John R.",
            "Paul, Rand", "Perry, James R. (Rick)",
            "Rubio, Marco", "Trump, Donald J.",
            "Walker, Scott" )
attach(az)
az$party[cand_nm %in% index] <- "independent"
az$party[cand_nm %in% dindex] <- "democrat"
az$party[cand_nm %in% rindex] <- 'republican'
detach(az)

# Convert party to factor
az$party <- factor(az$party)
```

I also would like to add other information such as latitudes and longitudes for map analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}
data(zipcode)
az_loc <- subset(zipcode, state == 'AZ')
az$clean_zip <- substring(az$contbr_zip, 1, 5)
az <- merge(az, az_loc, by.x = 'clean_zip', by.y = 'zip')




```

I would also like to integrate population data by zip-code from the 2010 ZCTA census.



```{r echo=FALSE, message=FALSE, warning=FALSE}


# Add Gender col to candidates

# Gender indices
m_index <- c('Bush, Jeb', "Carson, Benjamin S."
            , "Christie, Christopher J", "Cruz, Rafael Edward 'Ted'",
            "Gilmore, James S III" ,
            "Graham, Lindsey O.", "Huckabee, Mike", 
            "Jindal, Bobby", "Kasich, John R.",
            "Paul, Rand", "Perry, James R. (Rick)",
            "Rubio, Marco", "Trump, Donald J.",
            "Walker, Scott", "Sanders, Bernard",
            "Lessig, Lawrence", "O'Malley, Martin Joseph",
            "Webb, James Henry Jr.", "Johnson, Gary",
            "McMullin, Evan"
            )

f_index <- c("Clinton, Hillary Rodham", "Fiorina, Carly", "Stein, Jill" )


head(az)
#simple cand_gender
az$cand_gender <-'a'
attach(az)
az$cand_gender[cand_nm %in% m_index] <- "Male"
az$cand_gender[cand_nm %in% f_index] <- "Female"
detach(az)
# convert cand_gender to factor
az$cand_gender <- factor(az$cand_gender)

az$cand_gender

az$cand_nm %in% m_index

az

```
Now that I added candidates' genders, I'll add the contributors' genders, by using the gender package.
```{r echo=FALSE, message=FALSE, warning=FALSE}
# contributers' genders



# Get first names in a seperate col
az$first_name <- str_split_fixed(az$contbr_nm, ", ", 2)[,2]

# Use gender function
gender_df <- gender(as.character(az$first_name), c(1932, 1998),
countries= "United States")

# Assign gender to contributers in az df
names(gender_df)[1] = "first_name"
names(gender_df)[4] = 'contrib_gender'
gender_df <- unique(gender_df)
az <- merge(az, gender_df[ c("first_name", "contrib_gender")])

# convert contrib_gender to factor
az$contrib_gender <- factor(az$contrib_gender)














```


### Exploratory Data Analysis
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a function that gets  top ten counts
top10 <- function(x){
  y <-table(x)
  y <- sort(y, decreasing = T)
  y <- as.data.frame(y)
  y[1:10,]

  
  
}
  
  
  

# count city occurences in data set
city_tab <- top10(az$contbr_city)


ggplot(city_tab, aes(x, Freq))+
  geom_bar(stat='identity')

# top contribs counts
top_parties <-top10(az$party)

ggplot(top_parties, aes(x, Freq))+
  geom_bar(stat='identity')




city_tab


```

I would like to know how the data is distributed. 


```{r echo=FALSE, message=FALSE, warning=FALSE}

#money disterbution

ggplot(az, aes(contb_receipt_amt))+
  geom_histogram(binwidth = 15)+
  coord_cartesian(xlim=c(0:500))


#money disterbution normalized


ggplot(az, aes(contb_receipt_amt))+
  geom_histogram()+
  scale_x_log10()






ggplot(az, aes(contb_receipt_amt))+
  geom_freqpoly()





```


Date distribution 
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Date disterbution
ggplot(az, aes(proper_date))+
  geom_histogram(binwidth = 20)






```

It appears that we have negative numbers, that goes all to -5400. I believe that it represents refunds, since the most receipt comment is receipt.

I want to find out the amount stats without the refunds.


```{r echo=FALSE, message=FALSE, warning=FALSE}

# most popular bill denomination/   
summary(subset(az$contb_receipt_amt, az$contb_receipt_amt > 0 ))
non_zero_rec <- subset(az$contb_receipt_amt, az$contb_receipt_amt > 0 )
tab <- table(non_zero_rec)
str(tab)


tab[tab>1000]


top_deno<- top10(az$contb_receipt_amt)

ggplot(top_deno, aes(x, Freq))+
  geom_bar(stat='identity')




```

I wonder why the odd numbers such as 19, 27 or even 38.  


Below we will see the number of contributions for each candidate and how they break out: 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Most count of cont.
sort(summary(az$cand_nm), decreasing= T)


# top  cand in terms of money
cand_groups <- group_by(az, cand_nm) 

cand_sum <-summarize(cand_groups, 
                      mean(contb_receipt_amt),
                      n = n())


tot_rec <- aggregate(az$contb_receipt_amt, list(az$cand_nm), sum)


tot_rec <- arrange(tot_rec, desc(x))

tot_rec

# Top in money recieved histogram

ggplot(data= tot_rec,
       aes(reorder(Group.1, -x), x)) +
  geom_bar(stat="identity")
  

# Normalized
ggplot(data= tot_rec,
       aes(reorder(Group.1, -x), x)) +
  geom_bar(stat="identity")+
  scale_y_log10()



```

I would like to see the box-plot of each gender/party contribution.
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(az[!is.na(az$party) & az$party != 'independent',],
       aes(x=contrib_gender, y=contb_receipt_amt))+
  geom_boxplot(varwidth=T)+
  facet_grid(~party)+
  coord_cartesian(ylim=boxplot.stats(az$contb_receipt_amt)$stats[c(1, 5)]
)




```
Republicans contributed more on average, and they had a higher range of contribution amounts. Male republicans contributed slightly more on average than their female counterparts. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
table(az$contrib_gender)
```
I did not expect to find more female contributors than males in this data-set. 

Lets explore if females were more likely to vote for female candidates.
```{r echo=FALSE, message=FALSE, warning=FALSE}
same_sexf<- subset(az, az$contrib_gender == 'female' &
                     az$cand_gender== 'Female')

length(same_sexf$contrib_gender)/
  length(az$contrib_gender[az$contrib_gender == 'female'])


same_sexm<- subset(az, az$contrib_gender == 'male' & az$cand_gender== 'Male')

length(same_sexm$contrib_gender)/
  length(az$contrib_gender[az$contrib_gender == 'male'])


ggplot(az[!is.na(az$contrib_gender) & !is.na(az$cand_gender),] ,
       aes(cand_gender, contrib_gender))+
  geom_bin2d()


```
%54.5 females of this data-set contributed to females, while %65.4 of males contributed to males, which is a negligible preference. 



```{r echo=FALSE, message=FALSE, warning=FALSE}

# Create the mode function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

# get distinct empoyee contrib from each employer

employer<- summary(az$contbr_employer)



                                  
# most genrous occupations and account of total 
                                  

#subset academia employees                               
academia <- subset(az, grepl('UNIVERSITY|COLLEGE', az$contbr_employer,
                             ignore.case = T))


# group by university/college
cand_acad <- group_by(academia, contbr_employer)

# how much is averagly contributed by academia
spend_acad <- summarise(cand_acad,
                        number_of_contributions = n(),
                        avg_spent = mean(contb_receipt_amt),
                        main_cand = getmode(cand_nm),
                        party = getmode(party)
                        
                        )

# sort by num of contribs
attach(spend_acad)
spend_acad <- spend_acad[order(-number_of_contributions),]
detach(spend_acad)
spend_acad

# percentage of demo/repub colleges
percent(spend_acad$party)

# Percentage of candidate preferance
sort(percent(academia$cand_nm)[1:24],decreasing = T)

# 

```
Around %80 of colleges had a democratic preference. The majority of %59 of contributions were for Clinton, Sanders cones in second of %37. Cruz came in third (%1.42) and Trump close fourth (%1.4).


Below I will find the stats of homemakers and retirees
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Homemaker stats
homemaker <- subset(az, grepl('HOMEMAKER', az$contbr_occupation, ignore.case = T ))
summary(homemaker)

# Percentage of homemaker genders
percent(homemaker$contrib_gender)

# Homemaker contribs party leaning
percent(homemaker$party)





# retired findings
retired <- subset(az, grepl('RETIRED', az$contbr_occupation, ignore.case = T))

#percentage of retirees in the data
print(
length(unique(retired$contbr_nm))/
  length(unique(az$contbr_nm)) * 100)



# Retired party leaning percentage
print (summary(retired$party)/ length(retired$party) *100)

# Retrired avg spending
mean(retired$contb_receipt_amt)










```
Homemakers are %96 females, and  %65 of homemakers are democrats.

As we can see above, retirees make up about %35.6 of the data-set. Around %60 of retirees contributed to democrats and around %40 percent to republicans, contributions to independents are negligible. Retirees contributed $81 on average.



I want to know which occupations are most politically active, and how do they lean politically.

```{r echo=FALSE, message=FALSE, warning=FALSE}
jobs <-data.matrix(summary(az$contbr_occupation))

head(jobs,10)
```
The most politically active occupations in the data set are attorneys, physicians then teachers.

Below I would like to know the proportions of party leaning for each job. For example, of all engineers how many percent of them lean republican (number of republican engineers/ total number of engineers).
```{r echo=FALSE, message=FALSE, warning=FALSE}
require(GGally)
require(gmodels)


# Create a DF for jobs and their party count
job_party_table<- as.data.frame.matrix(table(az$contbr_occupation, az$party))

# Create a DF with jobs and thier party percentages
job_party_dist<- prop.table(table(az$contbr_occupation, az$party), 1)

job_party_dist <- as.data.frame.table(job_party_dist)

# Merge the two DFs above
job_party_props<- data.frame(c(job_party_table, job_party_dist))

#Keep these cols
keep <- c('Var1','democrat', 'republican', 'Var2', 'Freq')
job_party_props <- job_party_props[,keep]

#change props to percentages
job_party_props$Freq <- job_party_props$Freq *100
# add a total_job col
job_party_props$total <- job_party_props$democrat + job_party_props$republican

#drop these cols
job_party_props$democrat <- NULL
job_party_props$republican <- NULL

# change Var2 col to party
names(job_party_props)[names(job_party_props) == 'Var2'] <- 'party'


# subset top 10 republican jobs
top10rep <-subset(job_party_props, party == 'republican' & total > 1000)

# create a pie chart
pie(top10rep$Freq, labels = top10rep$Var1)

#create a bar plot
ggplot(top10rep,
       aes(Var1, Freq))+
  geom_bar(stat = "identity")



# subset top 10 democrats jobs
top10demo <-subset(job_party_props, party == 'democrat' & total > 1000
                   & quantile(job_party_props$Freq, c(.660), na.rm = T))

# create a bar plot
ggplot(top10demo,
       aes(Var1, Freq))+
  geom_bar(stat = "identity")


# subset top 10 independents jobs
top10inde <-subset(job_party_props, party == 'independent' & 
                     Freq > quantile(Freq, probs = .62, na.rm = T)
                   & total >100 & Var1 != 'N/A')

# create a bar plot
ggplot(top10inde,
       aes(Var1, Freq))+
  geom_bar(stat = "identity")



```






I would like to see average spending along dates
```{r echo=FALSE, message=FALSE, warning=FALSE}
#group by avg week
require(lubridate)

by_wk_sum <- tapply(az$contb_receipt_amt, week(az$proper_date), sum)

by_wk_avg <- tapply(az$contb_receipt_amt, week(az$proper_date), mean)


plot(by_wk_avg, type = 'l')

plot(by_wk_sum)



# note the weeks are aggregated by all years
az$week <-format(az$proper_date, format = "%W")
az$month <-format(az$proper_date, format = "%m")
az$year <-format(az$proper_date, format = "%y")

by_wk <- az %>% group_by(year = as.numeric(year), week= as.numeric(week), party) %>% summarise(sum = sum(contb_receipt_amt), 
              avg = mean(contb_receipt_amt),
              n=n())



ggplot(subset(by_wk, avg> 0 & year > 14),
       aes(as.numeric(week), avg, color=party, size=n))+ geom_line()+
  facet_grid(~year)


```

It seems that the avg amount of contributions are huge at the beginning of 2015, but when I added a 4th variable (n = number of contributions) it shows that these were a few outliers, the mass of the contributions came in mid 2016 as it lowered the average but the size (n) was bigger substantially. 








